Supporting Information

### 2015-12-14 ms.Rmd
Added our in situ results (related to #22)

At least `r round(min(st_libStats$"Reads (M)"), 0)` million single-end, 50-base reads were produced from each library, yielding between `r round(min(st_libStats$"Reads in genes (M)", na.rm = TRUE), 0)` and `r round(max(st_libStats$"Reads in genes (M)", na.rm = TRUE), 0)` million reads that mapped to genes after excluding reads likely to have originated from rRNA and tRNA. Using strict cut-offs for gene expression, `r length(unique(readRDS("output/expressedGenes/expressedGenesAll.Rds")))` unique genes were detected in two or more libraries from at least one stage. As a preliminary assessment of the RNA sequencing (RNAseq) results, the qualitative and quantitative expression of `r nIS` genes were compared with patterns that had previously been reported. `r nPos` of the `r nIS` genes were detected in two or more libraries from at least one meristem type in the LMD dataset (`r I(pasteLabel("Figure S", s_figCount, "sf_isGenesTpm"))` and `r I(pasteLabel("Table S", s_tableCount, "st_refTable"))`). Four genes that were detected in specific meristem samples were selected for validation by *in situ* hybridization (`r I(pasteLabel("Figure S", s_figCount, "sf_inSitu"))`).

@@

**`r I(pasteLabel("Figure S", s_figCount, "sf_inSitu"))`**: Placeholder for the *in situ* results figure.

>>>>

At least `r round(min(st_libStats$"Reads (M)"), 0)` million single-end, 50-base reads were produced from each library, yielding between `r round(min(st_libStats$"Reads in genes (M)", na.rm = TRUE), 0)` and `r round(max(st_libStats$"Reads in genes (M)", na.rm = TRUE), 0)` million reads that mapped to genes after excluding reads likely to have originated from rRNA and tRNA. Using strict cut-offs for gene expression, `r length(unique(readRDS("output/expressedGenes/expressedGenesAll.Rds")))` unique genes were detected in two or more libraries from at least one stage. As a preliminary assessment of the RNA sequencing (RNAseq) results, the qualitative and quantitative expression of `r nIS` genes were compared with patterns that had previously been reported. `r nPos` of the `r nIS` genes were detected in two or more libraries from at least one meristem type in the LMD dataset (`r I(pasteLabel("Figure S", s_figCount, "sf_isGenesTpm"))` and `r I(pasteLabel("Table S", s_tableCount, "st_refTable"))`). The expression of further four genes that were detected in specific meristem types by RNAseq was confirmed by RNA *in situ* hybridization. Each of the four genes was detected in the same meristem types by both methods (`r I(pasteLabel("Figure S", s_figCount, "sf_inSitu"))`).

@@

**`r I(pasteLabel("Figure S", s_figCount, "sf_inSitu"))`**: RNA *in situ* hybridization analysis to confirm the specific expression patterns of four genes detected in the RNA sequencing dataset.

### 2015-12-14 si.Rmd
Added our in situ results (related to #22)

\begin{figure*}
 \includegraphics{data/lmdFigure.pdf}
\caption{Placeholder for the \emph{in situ} results figure.}
\end{figure*}

@@

The primary sequences of *Arabidopsis thaliana*, *Solanum lycopersicum* and *Oryza sativa* MADS proteins were downloaded by querying Phytozome 10.3 [@Goodstein:2012ho] for the PANTHER family PTHR11945 ('MADS BOX PROTEIN') via the `biomaRt` interface [@Thomas:2003db; @Durinck:2009ki]. An initial alignment of all proteins was performed with `Clustal Omega` using a full distance matrix [@Sievers:2011fn]. To clean the alignment, poorly conserved regions where greater than `r minpcnongap` percent of sequences had gaps or less than `r minpcident` percent of pairwise comparisons between sequences had identical residues were removed. To recover MADS genes related to the Arabidopsis MIKC clade, a UGPMA tree was estimated from the cleaned alignment and the tree was cut at the lowest height that returned a clade including all type II Arabidopsis MADS proteins listed in the Arabidopsis transcription factor database on the Arabidopsis Gene Regulatory Information Server [@Yilmaz:2011ew], with as few type I proteins as possible. Protein sequences in this clade that were longer than `r minProtLength` amino acids were realigned using `Clustal Omega`, and the alignment was cleaned as described above and used to estimate a neighbour-joining tree with the `bionj` algorithm implemented in the `ape` package [@Paradis:2004dv] rooted to type I MADS sequences (`r og`). L~2~FC change expression values for tomato and Arabidopsis genes were estimated from samples that have previously been described [@Mantegazza:2014di; @Park:2012dg] using the same method described above for the rice samples. The tree was plotted using the `ggtree` extension for `ggplot2` ([@Wickham:2009tn]; https://github.com/GuangchuangYu/ggtree).

>>>>

\clearpage
\begin{figure*}
  \includegraphics{data/isFigure.pdf}
\end{figure*}
\captionof{figure}{RNA \emph{in situ} hybridization analysis to confirm the specific expression patterns of four genes detected in the RNA sequencing dataset. Expression of \emph{LOC\_Os09g27730} (a--d), \emph{LOC\_Os01g04670} (e--h), \emph{LOC\_Os10g04270} (i--l) and \emph{LOC\_Os10g05990} (m--p) was analysed at the rachis meristem (a, e, i and m), primary branch meristem (b, f, j and n), elongating primary branch meristem plus axillary meristem (c, g, k and o) and spikelet meristem (d, h, l and p) stages. Each gene was detected in the same stage by \emph{in situ} hybridization as by RNA sequencing.}

@@

The primary sequences of *Arabidopsis thaliana*, *Solanum lycopersicum* and *Oryza sativa* MADS proteins were downloaded by querying Phytozome 10.3 [@Goodstein:2012ho] for the PANTHER family PTHR11945 ('MADS BOX PROTEIN') via the `biomaRt` interface [@Thomas:2003db; @Durinck:2009ki]. An initial alignment of all proteins was performed with `Clustal Omega` using a full distance matrix [@Sievers:2011fn]. To clean the alignment, poorly conserved regions where greater than `r minpcnongap` percent of sequences had gaps or less than `r minpcident` percent of pairwise comparisons between sequences had identical residues were removed. To recover MADS genes related to the Arabidopsis MIKC clade, a UGPMA tree was estimated from the cleaned alignment and the tree was cut at the lowest height that returned a clade including all type II Arabidopsis MADS proteins listed in the Arabidopsis transcription factor database on the Arabidopsis Gene Regulatory Information Server [@Yilmaz:2011ew], with as few type I proteins as possible. Protein sequences in this clade that were longer than `r minProtLength` amino acids were realigned using `Clustal Omega`, and the alignment was cleaned as described above and used to estimate a neighbour-joining tree with the `bionj` algorithm implemented in the `ape` package [@Paradis:2004dv] rooted to type I MADS sequences (`r og`). L~2~FC change expression values for tomato and Arabidopsis genes were estimated from samples that have previously been described [@Mantegazza:2014di; @Park:2012dg] using the same method described above for the rice samples. The tree was plotted using the `ggtree` extension for `ggplot2` [https://github.com/GuangchuangYu/ggtree; @Wickham:2009tn].
